---
title: "The Action Dynamics of Native and Non-Native Speakers of English in Processing Active and Passive Sentences"
author: "Nicholas Duran"
date: 01/09/18
output: html_document
---

# Analysis 1: Interaction Between Language Proficiency and Grammatical Construction

In this analysis, we used linear mixed effects modeling to evaluate the fixed effect factors for grammatical structure (Active; coded as 0.5 vs. Passive; coded as -0.5) and language experience (Native; coded as 0.5 vs. Non-native; coded as -0.5) for each of the trajectory properties. We also generate a correlation matrix between trajectory properties and report descriptives.  

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
setwd('/Users/nduran/Dropbox (ASU)/ScottProject/activePassive/dataAnalysis2/')

library(dtplyr)
library(dplyr) # NEW - for data prep 
library(reshape2) # for melting data, for plotting
library(lme4) # mixed effects modeling
library(multcomp) # to do multiple pairwise comparisons when have multiple levels to a categorical predictor
library(MuMIn) # compute effect size (for mixed effects modeling)
library(openxlsx) # new way to save to excel
library(Hmisc)
library(data.table) # easiest way to rename a bunch of variables
library(pander)

source("save_model.r")

```

## Preliminaries

**Clean-up (dropping errant data):**

```{r}
combDat = read.delim("PROCESSED_SUPERNEW.txt", header = TRUE,sep='\t') ## data for non-native speakers

## remove practice trials
combDat = filter(combDat,condition != "practice") 

## drop participants where more than 30% of trials are considered errors (74 total trials)
badSubs = dplyr::summarize(group_by(combDat,subjectN,subjectOG,ability),
                         count=(sum(error)/length(subjectN))*100)
bad1 = filter(badSubs, count > 30)
combDat2 = filter(combDat, !(subjectN %in% bad1$subjectN))

## dropping participants based on visual inspection of trajectories that were highly unusual
combDat3 = filter(combDat2, !(subjectOG %in% c(277,2005,2776,3472,2649,3090,678, 2957)))

## drop all remaining trials with errors
combDat4 = filter(combDat3, error != 1) 

## generate clean data set - remove all but the declarative and passive types
combDat5 = filter(combDat4,condition == "passive" | condition == "declarative")

```

**Recoding to make labels more intuitive:**

```{r}

## rename variables to be more intuitive
setnames(combDat5, old = c('rt','initTime','DV_inmot','DV_velmax','DV_velmax_start','DV_dist','DV_xflp_tot','DV_AUC'), new = c('Overall Time','Latency Time','Motion Time','Velocity (Max)','Velocity (Onset)','Distance','X Flips','AUC'))

## rename variables to be more intuitive; df1 to be used for generating correlations
df1 <- within( combDat5, {
    subject.id <- subjectN ## subject number
    item.id <- stim ## item code
    
    ## Factor 1: Proficiency based on NS or NNS (2-levels; between-subject)
    proficiency <- ifelse( ability==1,"NS",
                    ifelse( ability==2,"NNS",NA ))
    ## Factor 2: Constructions of either Passive vs. Active (2-levels; within-subject)
    construction <- ifelse( condition=="declarative","Active",
                    ifelse( condition=="passive","Passive",NA ))

} )

#### NEED TO UPDATE
df1 <- dplyr::select(df1, subject.id,item.id,proficiency,construction,c(16:21,23:24)) 

## rename variables to be more intuitive; df2 to be used for descriptives/statistical analyses

#### NEED TO UPDATE
df2 = melt(df1,id=1:4) 
df2 <- within( df2, {
    dependent.variable <- value ## action dynamic response variables
    traj.properties <- variable
} )

```

**Descriptives of main variables:**

```{r}
summt = dplyr::summarize(group_by(df2,proficiency,construction,traj.properties),
                         count=length(dependent.variable),
                         mean=mean(dependent.variable),
                         se=sd(dependent.variable)/sqrt(length(dependent.variable)),
                         sdev=sd(dependent.variable))
# write.xlsx(summt,"descriptivesActive.xlsx")
# pander(summt[c(1:6)])

```

```{r}

## REVIEWER REQUEST - calculating confidence intervals from a normal distribution
# http://onlinestatbook.com/2/estimation/difference_means.html

makeList = c()
DVS = c('Overall Time','Latency Time','Motion Time','Velocity (Max)','Velocity (Onset)','Distance','X Flips','AUC')
getNS = CIfunc(DVS,summt,"NS")
getNNS = CIfunc(DVS,summt,"NNS")
makeList[[1]] = getNS
makeList[[2]] = getNNS
names(makeList) <- c("NS", "NNS")
# write.xlsx(makeList,"getCI_2.xlsx",rowNames=TRUE)
# makeList
```

## Correlations 

To examine the extent to which measures are separable or index shared aspects of processing

```{r}

#### NEED TO UPDATE

rcorMouse3 = rcorr(as.matrix(df1[,c(5:12)], type="pearson"))
# rcorMouse3[1]
# write.xlsx(rcorMouse3[1],"mouseCorrelations.xlsx")
```

## Statistical Models

**Contrast structure:**

```{r}
df3 <- within( df2, {
    ##// actives compared to passives
    construction.c <- ifelse( construction=="Active", 1/2, ifelse( construction=="Passive", -1/2, NA ) ) 
    ##// NS compared to NNS
    proficiency.c <- ifelse( proficiency=="NS", 1/2, ifelse( proficiency=="NNS", -1/2, NA ) )
} )
```

**Run models (except x-flips):**

```{r}

model.summary = c()
effect.size = c()
model.interaction = c()
model.contrasts = c()
i = 1

DVS = c('Overall Time','Latency Time','Motion Time','Velocity (Max)','Velocity (Onset)','Distance','AUC')
# DVS = c('X Flips')

for (vari in DVS) {
    df3.stats = filter(df3, variable == vari )
    
    # **Run Full Model:**
    omnibus.lmer <- lmer( dependent.variable ~ construction.c * proficiency.c
            + (1 + construction.c | subject.id) 
            + (0 + construction.c | item.id),
            control = lmerControl(optimizer="bobyqa"), data=df3.stats, REML=FALSE, na.action = na.exclude)

    ## **Get effect size:**
    getES = runEffectSize(omnibus.lmer, vari)
    effect.size = cbind(effect.size,getES)
    
    ## **Test for relevant two-way interaction:**
    twoway.omnibus.lmer = update(omnibus.lmer,.~.-construction.c:proficiency.c)
    model.interaction[[i]] = modComp(omnibus.lmer,twoway.omnibus.lmer,vari)

    # **Interpreting the omnibus coefficients:**
    planCont = rbind(
        ## main effect 1
        "AvP" = c(0, 1, 0, 0),
        ## main effect 2
        "NSvNNS" = c(0, 0, 1, 0),
        ## 2:way interaction
        "construction:proficiency" = c(0, 0, 0, 1))
    model.contrasts[[i]] = runContrasts(omnibus.lmer, planCont, vari)
    i = i + 1    
    
}    

## run follow-up post-hoc analysis of the interaction involving AUC

NS.only = filter(df3, proficiency == "NS" )
lmer.NS <- lmer( dependent.variable ~ construction.c
                + (1 + construction.c | subject.id)
                + (0 + construction.c | item.id),
                control = lmerControl(optimizer="bobyqa"), data=NS.only, REML=FALSE, na.action = na.exclude,subset=variable=="AUC")
# summary(lmer.NS)
coefs = data.frame(summary(lmer.NS)$coef)
coefs$p = 2*(1-pnorm(abs(coefs$t.value)))   

NNS.only = filter(df3, proficiency == "NNS" )
lmer.NNS <- lmer( dependent.variable ~ construction.c
                + (1 + construction.c | subject.id)
                + (0 + construction.c | item.id),
                control = lmerControl(optimizer="bobyqa"), data=NNS.only, REML=FALSE, na.action = na.exclude,subset=variable=="AUC")
# summary(lmer.NNS)
coefs = data.frame(summary(lmer.NNS)$coef)
coefs$p = 2*(1-pnorm(abs(coefs$t.value)))   

# write.xlsx(model.summary,"modelResults1.xlsx",rowNames=TRUE)
# write.xlsx(effect.size,"modelResults_effectsizes.xlsx",rowNames=TRUE)
# write.xlsx(model.interaction,"modelResults3.xlsx",rowNames=TRUE)
# write.xlsx(model.contrasts,"modelResults_6DVs.xlsx",rowNames=TRUE)

```

**Run models (x-flips):**

```{r}

omnibus.lmer.DC <- glmer( dependent.variable ~ construction.c * proficiency.c 
                          + (1 + construction.c | subject.id)
                          + (0 + construction.c | item.id),
                          data=df3, family = poisson(link = "log"), na.action = na.exclude, 
                          subset=variable=="X Flips")

contrasts.DC = runContrasts(omnibus.lmer.DC, planCont, "X Flips")    
# contrasts.DC
# write.xlsx(contrasts.DC,"modelResults_XFLIPS.xlsx",rowNames=TRUE)

```

# Analysis 2: Non-native speakers with proficiency now based on TOEFL scores

In this analysis we examine non-native speakers only, with fixed effects for grammatical structure and language proficiency, coded as a mean-centered continuous variable of non-native speakers’ TOEFL scores.

## Preliminaries

**Recoding:**

```{r}

## add NNS TOEFL scores - add to "combDat" generated in Analysis 1
toef = read.delim("TOEFLSCORES.csv", header = TRUE,sep=',') 
toef$subjectOG = toef$Participant
toef2 = dplyr::select(toef,subjectOG,TOEFL.scores)

combDat6 = merge(combDat5, toef2, by=c("subjectOG"), all.x=TRUE) # NA's match, drop all NNS without TOEFL score

#### NEED TO CHANGE

df1 <- dplyr::select(combDat6, c(1:3,7,5,26,16:21,23:24)) 
df2 = melt(df1,id=1:6) 

## rename variables to be more intuitive
df3 <- within( df2, {
    dependent.variable <- value ## action dynamic response variables
    traj.properties <- variable    
    
    subject.id <- subjectN ## subject number
    item.id <- stim ## item code
    
    ## Factor 1: Constructions of either Passive vs. Active (2-levels; within-subject)
    construction <- ifelse( condition=="declarative","Active",
                    ifelse( condition=="passive","Passive",NA ))

} )

## for TOEFL SCORES in models
center_scale = function(x) {
    scale(x, scale = FALSE)
}
```

**Contrast structure:**

```{r}
df4 <- within( df3, {
    ##// actives compared to passives
    construction.c <- ifelse( construction=="Active", 1/2, ifelse( construction=="Passive", -1/2, NA ) ) 
} )
```

**Run models**

```{r}
effect.size = c()
model.contrasts = c()
i = 1

DVS = c('Overall Time','Latency Time','Motion Time','Velocity (Max)','Velocity (Onset)','Distance','X Flips','AUC')

for (vari in DVS) {
    df3.stats = filter(df4, variable == vari & ability==2)

    # **Run Full Model:**
    omnibus.lmer <- lmer( dependent.variable ~ construction.c*center_scale(TOEFL.scores)
                          + (1 + construction.c | subject.id)
                          + (0 + construction.c | item.id),
                          control = lmerControl(optimizer="bobyqa"), data=df3.stats, REML=FALSE, na.action = na.exclude)

    ## **Get effect size:**
    getES = runEffectSize(omnibus.lmer, vari)
    effect.size = cbind(effect.size,getES)

    # **Interpreting the omnibus coefficients:**
    planCont = rbind(
        ## main effect 1
        "AvP" = c(0, 1, 0, 0),
        ## main effect 2
        "TOEFL" = c(0, 0, 1, 0),
        ## 2:way interaction
        "Interaction" = c(0, 0, 0, 1))
    model.contrasts[[i]] = runContrasts(omnibus.lmer, planCont, vari)
    i = i + 1

}

names(model.contrasts) <- c('Overall Time','Latency Time','Motion Time','Velocity (Max)','Velocity (Onset)','Distance','X Flips','AUC')
# model.contrasts
# write.xlsx(model.contrasts,"modelResults_TOEFL.xlsx",rowNames=TRUE)
```




